% !TEX root = /Users/us2009801/Documents/GitHub/LinearAlgebra/main.tex
\chapter{Class UNK - Monday, October 30\ts{rd}, 2017}
\section{\S 2.9 Unknown}

\begin{imp:defn}{Dimension}{} The dimension of a non-zero subspace H, denoted dim(H), is the number of vectors in any basis for H. The dimension of ${o}$ is defined to be zero.
\end{imp:defn}
\begin{ex}
$\{ \begin{bmatrix}
1\\
0\\
\end{bmatrix}, \begin{bmatrix}
0\\
1\\
\end{bmatrix} \} $
\end{ex}
\begin{ex}
How do we know?\\
$\begin{bmatrix}
1 & 0\\
0 & 1\\
\end{bmatrix}$
is in REF, there is a picot in every row and column, so $\{ \begin{bmatrix}
1\\
0\\
\end{bmatrix}, \begin{bmatrix}
0\\
1\\
\end{bmatrix} \} $ is a basis for $\mathbb{R}^2$
\end{ex}
\begin{note}
It can be shown that any basis of a subspace H must have the same number of vectors.
\end{note}
\begin{imp:defn}{Rank}{} The rank of A is dim(Col(A)).
\end{imp:defn}
Why do we care about rank?\\
\begin{align*}
\text{Col(A)} &= \text{span of all columns of A}\\
&= \text{set of all linear combinations of the columns of A}\\
&= {x_1a_1+...+x_na_n}\\
&=\text{Ax for any x in } \mathbb{R}^n \\
&= \text{All b that we can solve Ax=b for}\\
\text{rank(A)}&=\text{dimension of above space}\\
\end{align*}
\begin{ex}
Determine the rank(A), dim(null(A)), for $$\begin{bmatrix}
2 & 5 & -3 &-4 &8\\
4 & 7 & -4 & -3 & 9\\
6 & 9 & -5 & 2 & 4\\
0 & -9 & 6 & 5 & 6\\
\end{bmatrix} \rightarrow \begin{bmatrix}
2 & 5 & -3 & -4 & 8\\
0 & -3 & 2 & 5 & -7\\
0 & 0 & 0 & 4 & 6\\
0 & 0 & 0 & 0 & 0\\
\end{bmatrix}$$
Rank(A)= number of pivot=3 columns\\
dim(nul(A)) =number of free = 2 variables
\end{ex}
\begin{imp:thm}{The Rank Theorem}{} If Matrix A has n columns, then Rank (A) + Dim(Nul(A))=n
\begin{enumerate}
\item Rank(A) = Number of pivot columns
\item Dim(Nul(A)) = Number of free variables
\item N= Total number of columns
\end{enumerate}
\end{imp:thm}
\begin{prob}
Suppose that a basis for Col(A) is $\{ \begin{bmatrix}
1\\
2\\
4\\
\end{bmatrix},\begin{bmatrix}
0\\
1\\
1\\
\end{bmatrix}\begin{bmatrix}
2\\
7\\
14\\
\end{bmatrix} \} $ and Nul(A) = $x_4 \begin{bmatrix}
2\\
-3\\
0\\
1\\
\end{bmatrix}$
\begin{subprob} What is dim(Col(A))? 3
\end{subprob}
\begin{subprob} What is dim(Nul(A))? 1
\end{subprob}
\begin{subprob} What is the rank of A? 3
\end{subprob}
\begin{subprob} What size matrix should A be? 3x4
\end{subprob}
\begin{subprob} What is the reduced row echelon form of A?
\begin{align*}
x_1&= 2x_4\\
x_2&=-3x_4\\
x_3&=0\\
x_4&=x_4\\
&=\\
x_1-2x_4&=0\\
x_2+3x_4&=0\\
x_3&=0
\end{align*}
$$\begin{bmatrix}
1 & 0 & 0 & -2\\
0 & 1 & 0 & 3\\
0 & 0 & 1 & 0\\
\end{bmatrix}$$
\end{subprob}
\begin{subprob} Find a matrix A by row reducing the matrix $$\begin{bmatrix}
1 & 0 & 2 & a_1\\
2 & 1 & 7 & a_2\\
4 & 1 & 14 & a_3\\
\end{bmatrix}$$ to reduced row echelon form, and then choosing the constants $a_1,a_2, a_3$ such that the reduced row echelon matches the form from the previous problem. How were the first three columns chosen?\\
$$\begin{bmatrix}
1 & 0 & 2 & a_1\\
2 & 4 & 7 & a_2\\
4 & 1 & 14 & a_3\\
\end{bmatrix}$$
$$\begin{bmatrix}
1 & 0 & 2 & a_1\\
0 & 1 & 3 & -2a_1+a_2\\
0 & 1 & 6 & -4a_1+a_3\\
\end{bmatrix}$$
$$\begin{bmatrix}
1 & 0 & 2 & a_1\\
0 & 1 & 3 & -2a_1+a_2\\
0 & 0 & 1 & \frac{-2a_1-a_2+a_3}{3}\\
\end{bmatrix}$$
$$\begin{bmatrix}
1 & 0 & 2 & a_1\\
0 & 1 & 3 & -2a_1+a_2\\
0 & 0 & 1 & 0\\
\end{bmatrix}$$
$$\begin{bmatrix}
1 & 0 & 2 & -2\\
0 & 1 & 3 & 3\\
0 & 0 & 1 & 0\\
\end{bmatrix}$$
\begin{align*}
a_1&=-2\\
-2a_1+a_2&=3\\
-2(-2)+a_2&=3\\
4+a_2&=3\\
a_2&=-1\\
&=\\
-2a_1-a_2+a_3&=0\\
-2(-2)-(-1)+a_3&=0\\
5+a_3&=0\\
a_3&=5\\
\end{align*}
\end{subprob}
\end{prob}
\begin{imp:thm}{The Basis Theorem}{} Let H be a p-dimensional subspace of $\mathbb{R}^n$. Any linearly independent set of p vectors in H is automatically a basis for H. Also, any set of p vectors in H that spans H is automatically a basis for H.
\end{imp:thm}
\begin{ex}
Could $\mathbb{R}^3$ possible contain a 4-dimensional subspace?\\
4-dimensional = 4 pivots in REF. But we only have 3 rows, so we can only have up to 3 pivots, so NO.
\end{ex}
\begin{ex}
What is the rank of a 3x4 matrix whose null space is three dimensional?\\
Rank is 1.
\end{ex}